rgbif: R client for working with GBIF species occurrence data

### 2015-11-14 manuscript.Rmd
update description about datasets, orgs, netowrks, inst, etc., thanks tim
added three use cases

@@

## Use cases

### A

### B

### C

@@

## Acknowledgements

@@

## Data Accessibility

>>>>

## GBIF feedback loop

With each request `rgbif` makes to GBIF's API, we send request headers that
tell GBIF that the request is coming from `rgbif`, including what version of
the package. This will help GBIF know what proportion of requests are coming from
this package, and therefore R, as most requests likely will come from `rgbif`.

@@

# Use cases

### Ecological niche modelling

```{r niche-modelling, child="use-cases/niche-modelling.Rmd"}
```

### Biodiversity in big cities

```{r city-diversity, child="use-cases/city-diversity.Rmd"}
```

### Valley oak occurrence data comparison

```{r oak-occurrences, child="use-cases/oak-occurrences.Rmd"}
```

@@

# Acknowledgements

@@

# Data Accessibility

### 2015-11-14 manuscript.tex
added three use cases

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textbf{{#1}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{longtable}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={rgbif: R client for working with GBIF species occurrence data},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

@@

Dataset metrics are another useful way of digging in to figure out what
datasets you may want to use. One drawback is that these data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (2474).

@@

\subsection{Use cases}\label{use-cases}

\subsubsection{A}\label{a}

\subsubsection{B}\label{b}

\subsubsection{C}\label{c}

@@

\subsection{Acknowledgements}\label{acknowledgements}

@@

\subsection{Data Accessibility}\label{data-accessibility}

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage{color}
\usepackage{fancyvrb}
\newcommand{\VerbBar}{|}
\newcommand{\VERB}{\Verb[commandchars=\\\{\}]}
\DefineVerbatimEnvironment{Highlighting}{Verbatim}{commandchars=\\\{\}}
% Add ',fontsize=\small' for more characters per line
\usepackage{framed}
\definecolor{shadecolor}{RGB}{248,248,248}
\newenvironment{Shaded}{\begin{snugshade}}{\end{snugshade}}
\newcommand{\KeywordTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{\textbf{{#1}}}}
\newcommand{\DataTypeTok}[1]{\textcolor[rgb]{0.13,0.29,0.53}{{#1}}}
\newcommand{\DecValTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\BaseNTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\FloatTok}[1]{\textcolor[rgb]{0.00,0.00,0.81}{{#1}}}
\newcommand{\CharTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\StringTok}[1]{\textcolor[rgb]{0.31,0.60,0.02}{{#1}}}
\newcommand{\CommentTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{\textit{{#1}}}}
\newcommand{\OtherTok}[1]{\textcolor[rgb]{0.56,0.35,0.01}{{#1}}}
\newcommand{\AlertTok}[1]{\textcolor[rgb]{0.94,0.16,0.16}{{#1}}}
\newcommand{\FunctionTok}[1]{\textcolor[rgb]{0.00,0.00,0.00}{{#1}}}
\newcommand{\RegionMarkerTok}[1]{{#1}}
\newcommand{\ErrorTok}[1]{\textbf{{#1}}}
\newcommand{\NormalTok}[1]{{#1}}
\usepackage{longtable}
\usepackage{graphicx}
% We will generate all images so they have a width \maxwidth. This means
% that they will get their normal width if they fit onto the page, but
% are scaled down if they would overflow the margins.
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth
\else\Gin@nat@width\fi}
\makeatother
\let\Oldincludegraphics\includegraphics
\renewcommand{\includegraphics}[1]{\Oldincludegraphics[width=\maxwidth]{#1}}
\ifxetex
  \usepackage[setpagesize=false, % page size defined by xetex
              unicode=false, % unicode breaks when used with xetex
              xetex]{hyperref}
\else
  \usepackage[unicode=true]{hyperref}
\fi
\hypersetup{breaklinks=true,
            bookmarks=true,
            pdfauthor={},
            pdftitle={rgbif: R client for working with GBIF species occurrence data},
            colorlinks=true,
            urlcolor=blue,
            linkcolor=magenta,
            pdfborder={0 0 0}}
\urlstyle{same}  % don't use monospace font for urls
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\setcounter{secnumdepth}{0}
% Pandoc toggle for numbering sections (defaults to be off)
\setcounter{secnumdepth}{0}
% Pandoc header

@@

\subsection{GBIF feedback loop}\label{gbif-feedback-loop}

With each request \texttt{rgbif} makes to GBIF's API, we send request
headers that tell GBIF that the request is coming from \texttt{rgbif},
including what version of the package. This will help GBIF know what
proportion of requests are coming from this package, and therefore R, as
most requests likely will come from \texttt{rgbif}.

@@

Dataset metrics are another useful way of digging in to figure out what
datasets you may want to use. One drawback is that these data are only
available for datasets of type \emph{checklist}, but there are quite a
lot of them (2483).

@@

\section{Use cases}\label{use-cases}

\subsubsection{Ecological niche
modelling}\label{ecological-niche-modelling}

In this example, we plot actual occurrence data for \emph{Bradypus}
species against a single predictor variable, BIO1 (annual mean
temperature). This is only ont step in a species distribution modelling
nworkflow.

This example can be done using BISON data as well with our rbison
package.

\emph{Load libraries}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"rgbif"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"dismo"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"maptools"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"plyr"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\emph{Raster files}

Make a list of files that are installed with the dismo package, then
create a rasterStack from these

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{files <-}\StringTok{ }\KeywordTok{list.files}\NormalTok{(}\KeywordTok{paste}\NormalTok{(}\KeywordTok{system.file}\NormalTok{(}\DataTypeTok{package =} \StringTok{"dismo"}\NormalTok{), }\StringTok{"/ex"}\NormalTok{, }\DataTypeTok{sep =} \StringTok{""}\NormalTok{),}
                    \StringTok{"grd"}\NormalTok{, }\DataTypeTok{full.names =} \OtherTok{TRUE}\NormalTok{)}
\NormalTok{predictors <-}\StringTok{ }\KeywordTok{stack}\NormalTok{(files)}
\end{Highlighting}
\end{Shaded}

\emph{Get world boundaries}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{data}\NormalTok{(wrld_simpl)}
\end{Highlighting}
\end{Shaded}

\emph{Get GBIF data using the rOpenSci package rgbif}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{nn <-}\StringTok{ }\KeywordTok{name_lookup}\NormalTok{(}\StringTok{"bradypus*"}\NormalTok{, }\DataTypeTok{rank =} \StringTok{"species"}\NormalTok{)}
\NormalTok{nn <-}\StringTok{ }\KeywordTok{na.omit}\NormalTok{(}\KeywordTok{unique}\NormalTok{(nn$data$nubKey))}
\NormalTok{df <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey =} \NormalTok{nn, }\DataTypeTok{hasCoordinate =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{limit =} \DecValTok{500}\NormalTok{)}
\NormalTok{df <-}\StringTok{ }\NormalTok{df[ }\KeywordTok{sapply}\NormalTok{(df, function(x) }\KeywordTok{class}\NormalTok{(x$data)) %in%}\StringTok{ "data.frame"} \NormalTok{]}
\NormalTok{df <-}\StringTok{ }\KeywordTok{ldply}\NormalTok{(}\KeywordTok{lapply}\NormalTok{(df, }\StringTok{"[["}\NormalTok{, }\StringTok{"data"}\NormalTok{))}
\NormalTok{df2 <-}\StringTok{ }\NormalTok{df[,}\KeywordTok{c}\NormalTok{(}\StringTok{'decimalLongitude'}\NormalTok{,}\StringTok{'decimalLatitude'}\NormalTok{)]}
\end{Highlighting}
\end{Shaded}

\emph{Plot}

\begin{enumerate}
\def\labelenumi{(\arabic{enumi})}
\itemsep1pt\parskip0pt\parsep0pt
\item
  Add raster data, (2) Add political boundaries, (3) Add the points
  (occurrences)
\end{enumerate}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(predictors, }\DecValTok{1}\NormalTok{)}
\KeywordTok{plot}\NormalTok{(wrld_simpl, }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{)}
\KeywordTok{points}\NormalTok{(df2, }\DataTypeTok{col =} \StringTok{"blue"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-19-1.pdf}
\caption{}
\end{figure}

\subsubsection{Biodiversity in big
cities}\label{biodiversity-in-big-cities}

In this example, we collect specimen records across different cities
using GBIF data from the \texttt{rgbif} package.

\emph{Load libraries}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{"rgbif"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"ggplot2"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"plyr"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"RCurl"}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{"RColorBrewer"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\emph{Get bounding boxes for some cites}

Bounding lat/long data is from
\href{https://raw.github.com/amyxzhang/boundingbox-cities/master/boundbox.txt}{here}.

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{rawdat <-}\StringTok{ }\KeywordTok{getURL}\NormalTok{(}\StringTok{'https://raw.githubusercontent.com/amyxzhang/boundingbox-cities/master/boundbox.txt'}\NormalTok{)}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{read.table}\NormalTok{(}\DataTypeTok{text =} \NormalTok{rawdat, }\DataTypeTok{header =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{sep=}\StringTok{"}\CharTok{\textbackslash{}t}\StringTok{"}\NormalTok{, }\DataTypeTok{col.names=}\KeywordTok{c}\NormalTok{(}\StringTok{"city"}\NormalTok{,}\StringTok{"minlat"}\NormalTok{,}\StringTok{"maxlon"}\NormalTok{,}\StringTok{"maxlat"}\NormalTok{,}\StringTok{"minlon"}\NormalTok{))}
\NormalTok{dat <-}\StringTok{ }\KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{city=}\NormalTok{dat$city, }\DataTypeTok{minlon=}\NormalTok{dat$minlon, }\DataTypeTok{minlat=}\NormalTok{dat$minlat, }\DataTypeTok{maxlon=}\NormalTok{dat$maxlon, }\DataTypeTok{maxlat=}\NormalTok{dat$maxlat)}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{getdata <-}\StringTok{ }\NormalTok{function(x)\{}
  \NormalTok{coords <-}\StringTok{ }\KeywordTok{as.numeric}\NormalTok{(x[}\KeywordTok{c}\NormalTok{(}\StringTok{'minlon'}\NormalTok{,}\StringTok{'minlat'}\NormalTok{,}\StringTok{'maxlon'}\NormalTok{,}\StringTok{'maxlat'}\NormalTok{)])}
  \NormalTok{num <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{geometry =} \NormalTok{coords)$meta$count}
  \KeywordTok{data.frame}\NormalTok{(}\DataTypeTok{city=}\NormalTok{x[}\StringTok{'city'}\NormalTok{], }\DataTypeTok{richness=}\NormalTok{num, }\DataTypeTok{stringsAsFactors =} \OtherTok{FALSE}\NormalTok{)}
\NormalTok{\}}
\end{Highlighting}
\end{Shaded}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{out <-}\StringTok{ }\KeywordTok{apply}\NormalTok{(dat, }\DecValTok{1}\NormalTok{, getdata)}
\end{Highlighting}
\end{Shaded}

\emph{Merge to original table}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{out <-}\StringTok{ }\KeywordTok{merge}\NormalTok{(dat, }\KeywordTok{ldply}\NormalTok{(out), }\DataTypeTok{by=}\StringTok{"city"}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\emph{Add centroids from bounding boxes}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{out <-}\StringTok{ }\KeywordTok{transform}\NormalTok{(out, }\DataTypeTok{lat =} \NormalTok{(minlat+maxlat)/}\DecValTok{2}\NormalTok{, }\DataTypeTok{lon =} \NormalTok{(minlon+maxlon)/}\DecValTok{2}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\emph{Plot data}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{mapp <-}\StringTok{ }\KeywordTok{map_data}\NormalTok{(}\StringTok{'world'}\NormalTok{)}
\KeywordTok{ggplot}\NormalTok{(mapp, }\KeywordTok{aes}\NormalTok{(long, lat)) +}
\StringTok{  }\KeywordTok{geom_polygon}\NormalTok{(}\KeywordTok{aes}\NormalTok{(}\DataTypeTok{group=}\NormalTok{group), }\DataTypeTok{fill=}\StringTok{"white"}\NormalTok{, }\DataTypeTok{alpha=}\DecValTok{0}\NormalTok{, }\DataTypeTok{color=}\StringTok{"black"}\NormalTok{, }\DataTypeTok{size=}\FloatTok{0.4}\NormalTok{) +}
\StringTok{  }\KeywordTok{geom_point}\NormalTok{(}\DataTypeTok{data=}\NormalTok{out, }\KeywordTok{aes}\NormalTok{(lon, lat, }\DataTypeTok{color=}\NormalTok{richness), }\DataTypeTok{size=}\DecValTok{5}\NormalTok{, }\DataTypeTok{alpha=}\FloatTok{0.8}\NormalTok{) +}
\StringTok{  }\KeywordTok{scale_color_continuous}\NormalTok{(}\DataTypeTok{low =} \StringTok{"#60E1EE"}\NormalTok{, }\DataTypeTok{high =} \StringTok{"#0404C8"}\NormalTok{) +}
\StringTok{  }\KeywordTok{labs}\NormalTok{(}\DataTypeTok{x=}\StringTok{""}\NormalTok{, }\DataTypeTok{y=}\StringTok{""}\NormalTok{) +}
\StringTok{  }\KeywordTok{theme_grey}\NormalTok{(}\DataTypeTok{base_size=}\DecValTok{14}\NormalTok{) +}
\StringTok{  }\KeywordTok{theme}\NormalTok{(}\DataTypeTok{legend.position =} \StringTok{"bottom"}\NormalTok{, }\DataTypeTok{legend.key =} \KeywordTok{element_blank}\NormalTok{()) +}
\StringTok{  }\KeywordTok{guides}\NormalTok{(}\DataTypeTok{color =} \KeywordTok{guide_legend}\NormalTok{(}\DataTypeTok{keywidth =} \DecValTok{2}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-24-1.pdf}
\caption{}
\end{figure}

\subsubsection{Valley oak occurrence data
comparison}\label{valley-oak-occurrence-data-comparison}

This example comes from \href{https://twitter.com/ajpelu}{Antonio J.
Perez-Luque} who
\href{https://twitter.com/ajpelu/status/473951167567757312}{shared his
plot on Twitter}. Antonio compared the occurrences of Valley Oak
(\emph{Quercus lobata}) from \href{http://www.gbif.org/}{GBIF} to the
distribution of the same species from the
\href{http://esp.cr.usgs.gov/data/little/}{Atlas of US Trees}.

\emph{Load libraries}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{library}\NormalTok{(}\StringTok{'rgbif'}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{'raster'}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{'sp'}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{'maptools'}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{'rgeos'}\NormalTok{)}
\KeywordTok{library}\NormalTok{(}\StringTok{'scales'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\emph{Get GBIF Data for Quercus lobata}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{keyQl <-}\StringTok{ }\KeywordTok{name_backbone}\NormalTok{(}\DataTypeTok{name=}\StringTok{'Quercus lobata'}\NormalTok{, }\DataTypeTok{kingdom=}\StringTok{'plants'}\NormalTok{)$speciesKey}
\NormalTok{dat.Ql <-}\StringTok{ }\KeywordTok{occ_search}\NormalTok{(}\DataTypeTok{taxonKey=}\NormalTok{keyQl, }\DataTypeTok{return=}\StringTok{'data'}\NormalTok{, }\DataTypeTok{limit=}\DecValTok{50000}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\emph{Get Distribution map of Q. lobata Atlas of US Trees (Little, E.)}

From
\href{http://esp.cr.usgs.gov/data/little/}{\url{http://esp.cr.usgs.gov/data/little/}}.
And save shapefile in same directory

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{url <-}\StringTok{ 'http://esp.cr.usgs.gov/data/little/querloba.zip'}
\NormalTok{tmp <-}\StringTok{ }\KeywordTok{tempdir}\NormalTok{()}
\KeywordTok{download.file}\NormalTok{(url, }\DataTypeTok{destfile =} \StringTok{"~/querloba.zip"}\NormalTok{)}
\KeywordTok{unzip}\NormalTok{(}\StringTok{"~/querloba.zip"}\NormalTok{, }\DataTypeTok{exdir =} \NormalTok{tmp)}
\NormalTok{ql <-}\StringTok{ }\KeywordTok{readShapePoly}\NormalTok{(}\KeywordTok{file.path}\NormalTok{(tmp, }\StringTok{"querloba.shp"}\NormalTok{))}
\end{Highlighting}
\end{Shaded}

\emph{Get Elevation data of US}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{alt.USA <-}\StringTok{ }\KeywordTok{getData}\NormalTok{(}\StringTok{'alt'}\NormalTok{, }\DataTypeTok{country =} \StringTok{'USA'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\emph{Create Hillshade of US}

\begin{Shaded}
\begin{Highlighting}[]
\NormalTok{alt.USA <-}\StringTok{ }\NormalTok{alt.USA[[}\DecValTok{1}\NormalTok{]]}
\NormalTok{slope.USA <-}\StringTok{ }\KeywordTok{terrain}\NormalTok{(alt.USA, }\DataTypeTok{opt =} \StringTok{'slope'}\NormalTok{)}
\NormalTok{aspect.USA <-}\StringTok{ }\KeywordTok{terrain}\NormalTok{(alt.USA, }\DataTypeTok{opt =} \StringTok{'aspect'}\NormalTok{)}
\NormalTok{hill.USA <-}\StringTok{ }\KeywordTok{hillShade}\NormalTok{(slope.USA, aspect.USA, }\DataTypeTok{angle =} \DecValTok{45}\NormalTok{, }\DataTypeTok{direction =} \DecValTok{315}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\emph{Plot map}

\begin{Shaded}
\begin{Highlighting}[]
\KeywordTok{plot}\NormalTok{(hill.USA, }\DataTypeTok{col =} \KeywordTok{grey}\NormalTok{(}\DecValTok{0}\NormalTok{:}\DecValTok{100}\NormalTok{/}\DecValTok{100}\NormalTok{), }\DataTypeTok{legend =} \OtherTok{FALSE}\NormalTok{, }\DataTypeTok{xlim =} \KeywordTok{c}\NormalTok{(-}\DecValTok{125}\NormalTok{, -}\DecValTok{116}\NormalTok{), }\DataTypeTok{ylim =} \KeywordTok{c}\NormalTok{(}\DecValTok{32}\NormalTok{, }\DecValTok{42}\NormalTok{), }\DataTypeTok{main =} \StringTok{'Distribution of Quercus lobata'}\NormalTok{, }\DataTypeTok{xlab =} \StringTok{"Long"}\NormalTok{, }\DataTypeTok{ylab =} \StringTok{'Lat'}\NormalTok{)}
\CommentTok{# add shape from Atlas of US Trees}
\KeywordTok{plot}\NormalTok{(ql, }\DataTypeTok{add =} \OtherTok{TRUE}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"white"}\NormalTok{, }\FloatTok{0.6}\NormalTok{), }\DataTypeTok{border =} \OtherTok{FALSE}\NormalTok{)}
\CommentTok{# add Gbif presence points}
\KeywordTok{points}\NormalTok{(dat.Ql$decimalLongitude, dat.Ql$decimalLatitude, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{7}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \KeywordTok{alpha}\NormalTok{(}\StringTok{"darkgreen"}\NormalTok{, }\FloatTok{0.8}\NormalTok{))}
\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{40.5}\NormalTok{, }\StringTok{"GBIF Data"}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{'darkgreen'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{, }\DataTypeTok{pt.cex =} \DecValTok{1}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{8}\NormalTok{)}
\KeywordTok{legend}\NormalTok{(}\DataTypeTok{x =} \NormalTok{-}\DecValTok{121}\NormalTok{, }\DataTypeTok{y =} \FloatTok{41.5}\NormalTok{, }\StringTok{"Atlas of United States Trees }\CharTok{\textbackslash{}n}\StringTok{ (Little, E. 1971)"}\NormalTok{, }\DataTypeTok{pt.cex =} \FloatTok{1.5}\NormalTok{, }\DataTypeTok{cex =} \NormalTok{.}\DecValTok{8}\NormalTok{, }\DataTypeTok{pch =} \DecValTok{19}\NormalTok{, }\DataTypeTok{col =} \StringTok{'white'}\NormalTok{, }\DataTypeTok{bty =} \StringTok{'n'}\NormalTok{)}
\end{Highlighting}
\end{Shaded}

\begin{figure}[htbp]
\centering
\includegraphics{components/figure/manuscript-unnamed-chunk-31-1.pdf}
\caption{}
\end{figure}

@@

\section{Acknowledgements}\label{acknowledgements}

@@

\section{Data Accessibility}\label{data-accessibility}
